
---
title: "Easy Test: ACF/PACF Comparison"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Objective

Compare standard `acf()` and `pacf()` from base R with `taperedacf()` from the `forecast` package on simulated AR/MA models.

```{r packages}
library(forecast)  # for taperedacf()
library(ggplot2)
library(gridExtra)
```

##Simulate an AR(2) Process

```{r ar2-sim}
set.seed(123)
ar2_series <- arima.sim(n = 200, model = list(ar = c(0.5, -0.3)))
ts.plot(ar2_series, main = "Simulated AR(2) Process")
```

##ACF Comparison

```{r acf-comparison}
acf_base <- acf(ar2_series, plot = FALSE)
acf_tapered <- taperedacf(ar2_series, type = "correlation", plot = FALSE)
acf_banded <- taperedacf(ar2_series, type = "correlation", banded = TRUE, plot = FALSE)

# Combine results into data frame for ggplot
acf_df <- data.frame(
  Lag = acf_base$lag,
  ACF_base = acf_base$acf,
  ACF_tapered = acf_tapered$acf,
  ACF_banded = acf_banded$acf
)

# Plot
ggplot(acf_df, aes(x = Lag)) +
  geom_line(aes(y = ACF_base, color = "Base R")) +
  geom_line(aes(y = ACF_tapered, color = "Tapered")) +
  geom_line(aes(y = ACF_banded, color = "Banded")) +
  labs(title = "ACF Comparison", y = "ACF", color = "Method") +
  theme_minimal()
```

##PACF Comparison

```{r pacf-comparison}
pacf_base <- pacf(ar2_series, plot = FALSE)
pacf_df <- data.frame(
  Lag = pacf_base$lag,
  PACF_base = pacf_base$acf
)

ggplot(pacf_df, aes(x = Lag, y = PACF_base)) +
  geom_segment(aes(xend = Lag, yend = 0), size = 1) +
  geom_hline(yintercept = 0, color = "gray") +
  labs(title = "PACF (Base R)", y = "PACF") +
  theme_minimal()
```

##Interpretation

- **Tapered ACF** smooths the tail of autocorrelations to reduce variance, especially useful for longer lags.
- **Banded ACF** zeroes out ACF beyond a certain lag (bandwidth), helping in sparse modeling.
- **Standard base R ACF/PACF** is unbiased but noisier at higher lags.

##Try on MA(1)

```{r ma1-sim}
ma1_series <- arima.sim(n = 200, model = list(ma = 0.8))
ts.plot(ma1_series, main = "Simulated MA(1) Process")

acf(ma1_series, main = "ACF of MA(1)")
taperedacf(ma1_series, main = "Tapered ACF of MA(1)")
```

